Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-05-05 paper.qmd
Added orange to high error
Started update the paper with the new version of quollr 0.2.6

distance_df_small_edges_two_curvy <- distance_two_curvy |>
    filter(distance < benchmark_two_curvy)

## Computed benchmark value
tr_from_to_df_two_curvy <- inner_join(
  tr_from_to_df_two_curvy, distance_df_small_edges_two_curvy,
  by = c("from", "to"))

@@

```{r}
#| label: langevitour-best-fit-two-curvy-proj-tsne
#| eval: false

@@

langevitour::langevitour(df_model_data_two_curvy_filtered[1:(length(df_model_data_two_curvy_filtered)-1)],
                         lineFrom = distance_df_small_edges_two_curvy$from,
                         lineTo = distance_df_small_edges_two_curvy$to,
                         group = factor(df_model_data_two_curvy_filtered$type,
                                        c("data", "model")),
                         levelColors = c(clr_choice, "#000000"))
```

```{r}
#| label: best-fit-two-curvy-proj-tsne
## First projection
model_prj1 <- cbind(
  c(0.37847,-0.06517,0.04231,0.01655),
  c(-0.06466,-0.27052,0.22131,-0.15235))

proj_obj1 <- get_projection(projection = model_prj1, 
               proj_scale = 1, 
               scaled_data = scaled_two_curvy_data, 
               scaled_data_model = scaled_two_curvy_data_model, 
               distance_df_small_edges = distance_df_small_edges_two_curvy, 
               axis_param = list(limits = 0.5, 
                                 axis_scaled = 2,
                                 axis_pos_x = -0.5, 
                                 axis_pos_y = -0.5, 
                                 threshold = 0.03))

@@

proj_obj2 <- get_projection(projection = model_prj2, 
                            proj_scale = 1, 
                            scaled_data = scaled_two_curvy_data, 
                            scaled_data_model = scaled_two_curvy_data_model, 
                            distance_df_small_edges = distance_df_small_edges_two_curvy, 
                            axis_param = list(limits = 0.35, 
                                              axis_scaled = 2,
                                              axis_pos_x = -0.35, 
                                              axis_pos_y = -0.35, 
                                              threshold = 0.01))

@@

```{r}
#| label: fig-two-curvy-true-proj
#| fig-cap: "Two \\gD{} projections from $4\\text{-}D$, for the two non-linear clusters data (blue points) and true model (black lines) are shown. The top cluster in (a) shows uniformly distributed data along a C-shape, while in the bottom cluster, the data is more concentrated in one corner of the curvy shape. Video of the langevitour animations are available at <>." 
#| fig-pos: H
#| fig-width: 10
#| fig-height: 5

@@

distance_two_curvy2 <- algo_obj_two_curvy2$distance_df
tr_from_to_df_two_curvy2 <- algo_obj_two_curvy2$tr_from_to_df
benchmark_two_curvy2 <- algo_obj_two_curvy2$benchmark
df_bin_centroids_two_curvy2 <- algo_obj_two_curvy2$df_bin_centroids
df_bin_two_curvy2 <- algo_obj_two_curvy2$df_bin
hex_grid_with_counts_two_curvy2 <- algo_obj_two_curvy2$hex_grid_with_counts

distance_df_small_edges_two_curvy2 <- distance_two_curvy2 |>
  filter(distance < benchmark_two_curvy2)

tr_from_to_df_two_curvy2 <- inner_join(
  tr_from_to_df_two_curvy2, distance_df_small_edges_two_curvy2, 
  by = c("from", "to"))

@@

distance_two_curvy3 <- algo_obj_two_curvy3$distance_df
tr_from_to_df_two_curvy3 <- algo_obj_two_curvy3$tr_from_to_df
benchmark_two_curvy3 <- algo_obj_two_curvy3$benchmark
df_bin_centroids_two_curvy3 <- algo_obj_two_curvy3$df_bin_centroids
df_bin_two_curvy3 <- algo_obj_two_curvy3$df_bin
hex_grid_with_counts_two_curvy3 <- algo_obj_two_curvy3$hex_grid_with_counts

distance_df_small_edges_two_curvy3 <- distance_two_curvy3 |>
  filter(distance < benchmark_two_curvy3)

tr_from_to_df_two_curvy3 <- inner_join(
  tr_from_to_df_two_curvy3, distance_df_small_edges_two_curvy3, 
  by = c("from", "to"))

@@

#nldr_two_curvy2 + 
  trimesh_two_curvy_removed1_tsne_with_data +
  two_curvy_proj_tsne_first_model1 +
  two_curvy_proj_tsne_all_model1 +
  plot_layout(ncol = 3)
```

@@

#### Removing long edges

Edges define the neighbourhood structure, in order to provide a smooth \gD{} representation of the fitted model. @fig-two-curvy-true-proj shows a wire frame of the true model that was used to generate the non-linear clusters example data. The ideal is that the representation of the fitted model, at least for this example where we know the true model, should look similar to this.

The Delaunay triangulation will ensure that all centroids are connected into a triangular mesh. For some structures, like clustered data, or highly non-linear shapes, breaks in the mesh are meaningful. When separated clusters are present the mesh should be broken across the gaps. For non-linear structures like the C-shaped curvilinear, the mesh should run unbroken along the C, but there should be no edges connecting the top of the C directly to the bottom of the C. For these reasons it is necessary to remove edges from the mesh in some applications.

The decision on edge length removal is made based on the distribution of edge lengths. In particular, a gap between values, where there a concentration of small values and then a few larger values, likely suggests a cutoff for edge removal. Because the triangulation is typically done on the hexagon centroids, there are particular discrete edge lengths, based on bin widths. @fig-rm-lg illustrates edge length distributions.

There is an additional step that is needed. When the model is lifted into \pD{}, if the fit is good all the edges should be relatively small in this space, too. If this is not the case, then there are several possible actions: (1) re-do the NLDR to get a more representative layout; (2) identify the edge and remove it from the model, in \gD{} and \pD{}; (3) consider different values for the model fit, number of bins, initial bin position or removing low density bins.

```{r}
#| label: model-two-curvy-umap

umap_two_curvy <- read_rds(file = "data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_umap_n-neigbors_15_min-dist_0.1.rds")

num_bins_x_two_curvy <- 42

algo_obj_two_curvy3 <- gen_nldr_vis_algo_obj(
  high_d_data = training_data_two_curvy, 
  nldr_data = umap_two_curvy, 
  num_x_bins = num_bins_x_two_curvy)

umap_two_curvy_scaled <- algo_obj_two_curvy3$nldr_scaled
distance_two_curvy <- algo_obj_two_curvy3$distance_df
tr_from_to_df_two_curvy <- algo_obj_two_curvy3$tr_from_to_df
benchmark_two_curvy <- algo_obj_two_curvy3$benchmark
df_bin_centroids_two_curvy <- algo_obj_two_curvy3$df_bin_centroids
df_bin_two_curvy <- algo_obj_two_curvy3$df_bin
hex_grid_with_counts <- algo_obj_two_curvy3$hex_grid_with_counts
a1 <- algo_obj_two_curvy3$a1

trimesh_two_curvy_umap <- ggplot() +
  geom_segment(data = tr_from_to_df_two_curvy,
               aes(
                 x = x_from,
                 y = y_from,
                 xend = x_to,
                 yend = y_to),
               colour = "#000000",
               alpha = 0.5,
               linewidth = 0.5) +
  geom_point(data = umap_two_curvy_scaled,
             aes(
               x = UMAP1,
               y = UMAP2
             ),
             color = "#636363",
             alpha = 0.5,
             size = 0.5
  ) +
  coord_equal() +
  xlim(c(0, 1)) +
  ylim(c(0, 1)) +
  interior_annotation("a1", sc_ltr_pos, cex = 2)

```{r}

# With the computed benchmark value
# distance_df_small_edges_two_curvy <- distance_two_curvy |>
#   filter(distance < benchmark_two_curvy)

####With diff benchmark values

## Benchmark 2

benchmark_two_curvy <- 8 * a1

distance_df_small_edges_two_curvy3 <- distance_two_curvy |>
  filter(distance < benchmark_two_curvy)

tr_from_to_df_two_curvy4 <- inner_join(
  tr_from_to_df_two_curvy, distance_df_small_edges_two_curvy3,
  by = c("from", "to"))

trimesh_two_curvy_removed2_umap <- ggplot() +
  geom_segment(data = tr_from_to_df_two_curvy4,
               aes(
                 x = x_from,
                 y = y_from,
                 xend = x_to,
                 yend = y_to),
               colour = "#000000",
               alpha = 0.5,
               linewidth = 0.5)+
  geom_point(data = umap_two_curvy_scaled,
             aes(
               x = UMAP1,
               y = UMAP2
             ),
             color = "#636363",
             alpha = 0.5,
             size = 0.5
  ) +
  coord_equal() +
  xlim(c(0, 1)) +
  ylim(c(0, 1)) +
  interior_annotation("b1", sc_ltr_pos, cex = 2)

benchmark_two_curvy <- 2.5 * bin_width

distance_df_small_edges_two_curvy2 <- distance_two_curvy |>
  filter(distance < benchmark_two_curvy) ## 0.231

tr_from_to_df_two_curvy3 <- inner_join(
  tr_from_to_df_two_curvy, distance_df_small_edges_two_curvy2,
  by = c("from", "to"))

trimesh_two_curvy_removed1_umap <- ggplot() +
  geom_segment(data = tr_from_to_df_two_curvy3,
               aes(
                 x = x_from,
                 y = y_from,
                 xend = x_to,
                 yend = y_to),
               colour = "#000000",
               alpha = 0.5,
               linewidth = 0.5)+
  geom_point(data = umap_two_curvy_scaled,
             aes(
               x = UMAP1,
               y = UMAP2
             ),
             color = "#636363",
             alpha = 0.5,
             size = 0.5
  ) +
  coord_equal() +
  xlim(c(0, 1)) +
  ylim(c(0, 1)) +
  interior_annotation("c1", sc_ltr_pos, cex = 2)
```

```{r}
#| label: compare-2d-and-4d-dist

### Compute highD distance
dist_vec <- proxy::dist(x = df_bin_two_curvy[, -1], method = "Euclidean") |> as.vector()

from_vec <- c()
to_vec <- c()
num_obs <- 1:(NROW(df_bin_two_curvy) - 1)

for (obs in num_obs) {

  from_val <- rep(obs, (NROW(df_bin_two_curvy) - obs))
  if ((obs + 1) <= NROW(df_bin_two_curvy)) {
    to_val <- (obs + 1):NROW(df_bin_two_curvy)
  }
  from_vec <- append(from_vec, from_val)
  to_vec <- append(to_vec, to_val)

}

dist_highd <- tibble::tibble(from = from_vec, to = to_vec, dist_highd = dist_vec)

## To plot the distribution of distances

# Define bin width and starting point
start_point <- a1/2

# Create bins and calculate the mean value for each bin range
bins <- seq(start_point, max(distance_two_curvy$distance) + a1,
            by = a1)
bin_labels <- bins[-length(bins)] + a1 / 2  # mean of each bin range

distance_two_curvy_dist <- distance_two_curvy |>
  mutate(bin = cut(distance, breaks = bins, include.lowest = TRUE, labels = bin_labels))

distance_two_curvy_dist <- left_join(distance_two_curvy_dist, dist_highd, by = c("from", "to"))

text_df <- tibble::tibble(
  x = c(0.12, 0.43),
  y = c(10, 10),
  text = c("c", "b")
)

distance_points_umap <- ggplot(
  distance_two_curvy_dist,
  aes(x = distance,
      y = dist_highd)) +
  geom_point(alpha = 0.5) +
  geom_text(
    data=text_df,
    aes(x=x, y=y,
        label=text),
    colour="#bdbdbd",
    size = 10) +
  geom_vline(xintercept = 8 * bin_width,
             linetype="solid",
             color = "#bdbdbd",
             linewidth=1) +
  geom_vline(xintercept = 2.5 * bin_width,
             linetype="solid",
             color = "#bdbdbd",
             linewidth=1) +
  ylab(expression(d^{(4)})) +
  xlab(expression(d^{(2)})) +
  theme_minimal() +
  theme(aspect.ratio = 0.75,
        panel.border = element_rect(fill = 'transparent'),
        plot.title = element_text(size = 20, hjust = 0.5, vjust = -0.5),
        panel.grid.major.x = element_blank(),
        axis.ticks.x = element_line(),
        axis.ticks.y = element_line(),
        axis.title = element_text(size = 15),
        axis.text = element_text(size = 12))
```

```{r}
#| label: proj-tsne-model

# Apply the scaling
df_model_data_two_curvy <- bind_rows(data_two_curvy, df_bin_two_curvy)

## First projection
projection <- cbind(
  c(0.37847,-0.06517,0.04231,0.01655),
  c(-0.06466,-0.27052,0.22131,-0.15235))

proj_obj1 <- get_projection(projection = projection, 
                            proj_scale = 1, 
                            scaled_data = scaled_two_curvy_data, 
                            scaled_data_model = scaled_two_curvy_data_model, 
                            distance_df_small_edges = distance_df_small_edges_two_curvy2, 
                            axis_param = list(limits = 0.6, 
                                              axis_scaled = 2,
                                              axis_pos_x = -0.5, 
                                              axis_pos_y = -0.5, 
                                              threshold = 0.04))

## Projection with Delauany triangulation

projected_model <- as.matrix(scaled_two_curvy_data_model) %*% projection_scaled

model_df1 <- dplyr::left_join(
  distance_two_curvy |> select(-distance),
  projected_model_df,
  by = c("from" = "ID"))

names(model_df1)[3:NCOL(model_df1)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")

model_df1 <- dplyr::left_join(model_df1, projected_model_df, by = c("to" = "ID"))
names(model_df1)[(2 + NCOL(projected_model_df)):NCOL(model_df1)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")

## With benchmark option 2

model_df2 <- dplyr::left_join(
  distance_df_small_edges_two_curvy3 |> select(-distance),
  projected_model_df,
  by = c("from" = "ID"))

names(model_df2)[3:NCOL(model_df2)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")

model_df2 <- dplyr::left_join(model_df2, projected_model_df, by = c("to" = "ID"))
names(model_df2)[(2 + NCOL(projected_model_df)):NCOL(model_df2)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")

two_curvy_proj_first_model1_umap <- plot_proj(
    proj_obj = proj_obj1, 
    point_param = c(0.5, 0.2, "#636363"), # size, alpha, color
    line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha
    plot_limits = c(-0.6, 0.6), 
    title = "c2", 
    cex = 2, 
    position = sc_ltr_pos,
    axis_text_size = 4, 
    is_color = FALSE) +
    theme(
        legend.position = "none"
    )

## Projection with Delauany triangulation

proj_obj1_temp1 <- proj_obj1
proj_obj1_temp2 <- proj_obj1
proj_obj1_temp1[["model_df"]] <- model_df1

two_curvy_proj_model_delaunay_umap <- plot_proj(
    proj_obj = proj_obj1_temp1, 
    point_param = c(0.5, 0.2, "#636363"), # size, alpha, color
    line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha
    plot_limits = c(-0.6, 0.6), 
    title = "a2", 
    cex = 2, 
    position = sc_ltr_pos,
    axis_text_size = 4, 
    is_color = FALSE) +
    theme(aspect.ratio = 1, 
          legend.position = "none")

proj_obj1_temp2[["model_df"]] <- model_df2

two_curvy_proj_model_benchmark2_umap <- plot_proj(
    proj_obj = proj_obj1_temp2, 
    point_param = c(0.5, 0.2, "#636363"), # size, alpha, color
    line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha
    plot_limits = c(-0.6, 0.6), 
    title = "b2", 
    cex = 2, 
    position = sc_ltr_pos,
    axis_text_size = 4, 
    is_color = FALSE) +
    theme(aspect.ratio = 1, 
          legend.position = "none")

```{r}
#| label: fig-rm-lg
#| fig-cap: "The choice of long edge removal significantly impacts the model's accuracy. \\gD{} Vs $7\\text{-}D$ distances between the bin centroids, where the model is fitted with UMAP for the non-linear clusters data. The two benchmarks for long edge removal: (b) benchmark = 8$a_1$ and (c) benchmark = 2.5$a_1$, where $a_1$ = 0.03 are made. The wireframes in \\gD{} and a projection from $7\\text{-}D$ are shown with Delaunay triangulation (a1, a2) and the two benchmark values (b1, b2, c1, c2). With the benchmark set at 6.5$a_1$, long edges exist within the C-shaped cluster, which distrupts the model. In contrast, using 2$a_1$ is a better option, as it effectively captures the non-linear shapes of the clusters. To generate an accurate model that captures the data structure, the choice of long edge removal does matter."
#| fig-height: 15
#| fig-width: 15

free(distance_points_umap) + 
    wrap_plots(trimesh_two_curvy_umap, trimesh_two_curvy_removed2_umap,
               trimesh_two_curvy_removed1_umap, two_curvy_proj_model_delaunay_umap,
               two_curvy_proj_model_benchmark2_umap, two_curvy_proj_first_model1_umap,
               ncol = 3) +
    plot_layout(heights = c(0.8, 1))
```

>>>>

@@

```{r}
#| label: langevitour-best-fit-two-curvy-proj-tsne

@@

langevitour::langevitour(df_model_data_two_curvy_filtered[1:(length(df_model_data_two_curvy_filtered)-1)],
                         lineFrom = tr_from_to_df_two_curvy$from,
                         lineTo = tr_from_to_df_two_curvy$to,
                         group = factor(df_model_data_two_curvy_filtered$type,
                                        c("data", "model")),
                         levelColors = c(clr_choice, "#000000"))
```

```{r}
#| label: best-fit-two-curvy-proj-tsne

## First projection
model_prj1 <- cbind(
  c(0.37847,-0.06517,0.04231,0.01655),
  c(-0.06466,-0.27052,0.22131,-0.15235))

proj_obj1 <- get_projection(projection = model_prj1, 
               proj_scale = 1, 
               scaled_data = scaled_two_curvy_data, 
               scaled_data_model = scaled_two_curvy_data_model, 
               distance_df_small_edges = tr_from_to_df_two_curvy, 
               axis_param = list(limits = 0.5, 
                                 axis_scaled = 2,
                                 axis_pos_x = -0.5, 
                                 axis_pos_y = -0.5, 
                                 threshold = 0.03))

@@

proj_obj2 <- get_projection(projection = model_prj2, 
                            proj_scale = 1, 
                            scaled_data = scaled_two_curvy_data, 
                            scaled_data_model = scaled_two_curvy_data_model, 
                            distance_df_small_edges = tr_from_to_df_two_curvy, 
                            axis_param = list(limits = 0.35, 
                                              axis_scaled = 2,
                                              axis_pos_x = -0.35, 
                                              axis_pos_y = -0.35, 
                                              threshold = 0.01))

@@

```{r}
#| label: fig-two-curvy-true-proj
#| fig-cap: "Two \\gD{} projections from $4\\text{-}D$, for the two non-linear clusters data (blue points) and true model (black lines) are shown. The top cluster in (a) shows uniformly distributed data along a C-shape, while in the bottom cluster, the data is more concentrated in one corner of the curvy shape. Video of the langevitour animations are available at <>." 
#| fig-pos: H
#| fig-width: 10
#| fig-height: 5
#| eval: false

@@

tr_from_to_df_two_curvy2 <- algo_obj_two_curvy2$tr_from_to_df
df_bin_centroids_two_curvy2 <- algo_obj_two_curvy2$df_bin_centroids
df_bin_two_curvy2 <- algo_obj_two_curvy2$df_bin
hex_grid_with_counts_two_curvy2 <- algo_obj_two_curvy2$hex_grid_with_counts

@@

tr_from_to_df_two_curvy3 <- algo_obj_two_curvy3$tr_from_to_df
df_bin_centroids_two_curvy3 <- algo_obj_two_curvy3$df_bin_centroids
df_bin_two_curvy3 <- algo_obj_two_curvy3$df_bin
hex_grid_with_counts_two_curvy3 <- algo_obj_two_curvy3$hex_grid_with_counts

@@

#nldr_two_curvy2 + 
  trimesh_two_curvy_removed1_tsne_with_data +
  #two_curvy_proj_tsne_first_model1 +
  #two_curvy_proj_tsne_all_model1 +
  plot_layout(ncol = 3)
```

@@

<!-- #### Removing long edges -->

<!-- Edges define the neighbourhood structure, in order to provide a smooth \gD{} representation of the fitted model. @fig-two-curvy-true-proj shows a wire frame of the true model that was used to generate the non-linear clusters example data. The ideal is that the representation of the fitted model, at least for this example where we know the true model, should look similar to this.  -->

<!-- The Delaunay triangulation will ensure that all centroids are connected into a triangular mesh. For some structures, like clustered data, or highly non-linear shapes, breaks in the mesh are meaningful. When separated clusters are present the mesh should be broken across the gaps. For non-linear structures like the C-shaped curvilinear, the mesh should run unbroken along the C, but there should be no edges connecting the top of the C directly to the bottom of the C. For these reasons it is necessary to remove edges from the mesh in some applications. -->

<!-- The decision on edge length removal is made based on the distribution of edge lengths. In particular, a gap between values, where there a concentration of small values and then a few larger values, likely suggests a cutoff for edge removal. Because the triangulation is typically done on the hexagon centroids, there are particular discrete edge lengths, based on bin widths. @fig-rm-lg illustrates edge length distributions.  -->

<!-- There is an additional step that is needed. When the model is lifted into \pD{}, if the fit is good all the edges should be relatively small in this space, too. If this is not the case, then there are several possible actions: (1) re-do the NLDR to get a more representative layout; (2) identify the edge and remove it from the model, in \gD{} and \pD{}; (3) consider different values for the model fit, number of bins, initial bin position or removing low density bins. -->

<!-- ```{r} -->
<!-- #| label: model-two-curvy-umap -->

<!-- umap_two_curvy <- read_rds(file = "data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_umap_n-neigbors_15_min-dist_0.1.rds") -->

<!-- num_bins_x_two_curvy <- 42 -->

<!-- algo_obj_two_curvy3 <- gen_nldr_vis_algo_obj( -->
<!--   high_d_data = training_data_two_curvy,  -->
<!--   nldr_data = umap_two_curvy,  -->
<!--   num_x_bins = num_bins_x_two_curvy) -->

<!-- umap_two_curvy_scaled <- algo_obj_two_curvy3$nldr_scaled -->
<!-- distance_two_curvy <- algo_obj_two_curvy3$distance_df -->
<!-- tr_from_to_df_two_curvy <- algo_obj_two_curvy3$tr_from_to_df -->
<!-- benchmark_two_curvy <- algo_obj_two_curvy3$benchmark -->
<!-- df_bin_centroids_two_curvy <- algo_obj_two_curvy3$df_bin_centroids -->
<!-- df_bin_two_curvy <- algo_obj_two_curvy3$df_bin -->
<!-- hex_grid_with_counts <- algo_obj_two_curvy3$hex_grid_with_counts -->
<!-- a1 <- algo_obj_two_curvy3$a1 -->

<!-- trimesh_two_curvy_umap <- ggplot() + -->
<!--   geom_segment(data = tr_from_to_df_two_curvy, -->
<!--                aes( -->
<!--                  x = x_from, -->
<!--                  y = y_from, -->
<!--                  xend = x_to, -->
<!--                  yend = y_to), -->
<!--                colour = "#000000", -->
<!--                alpha = 0.5, -->
<!--                linewidth = 0.5) + -->
<!--   geom_point(data = umap_two_curvy_scaled, -->
<!--              aes( -->
<!--                x = UMAP1, -->
<!--                y = UMAP2 -->
<!--              ), -->
<!--              color = "#636363", -->
<!--              alpha = 0.5, -->
<!--              size = 0.5 -->
<!--   ) + -->
<!--   coord_equal() + -->
<!--   xlim(c(0, 1)) + -->
<!--   ylim(c(0, 1)) + -->
<!--   interior_annotation("a1", sc_ltr_pos, cex = 2) -->

<!-- ``` -->

<!-- ```{r} -->

<!-- # With the computed benchmark value -->
<!-- # distance_df_small_edges_two_curvy <- distance_two_curvy |> -->
<!-- #   filter(distance < benchmark_two_curvy) -->

<!-- ####With diff benchmark values -->

<!-- ## Benchmark 2 -->

<!-- benchmark_two_curvy <- 8 * a1 -->

<!-- distance_df_small_edges_two_curvy3 <- distance_two_curvy |> -->
<!--   filter(distance < benchmark_two_curvy) -->

<!-- tr_from_to_df_two_curvy4 <- inner_join( -->
<!--   tr_from_to_df_two_curvy, distance_df_small_edges_two_curvy3, -->
<!--   by = c("from", "to")) -->

<!-- trimesh_two_curvy_removed2_umap <- ggplot() + -->
<!--   geom_segment(data = tr_from_to_df_two_curvy4, -->
<!--                aes( -->
<!--                  x = x_from, -->
<!--                  y = y_from, -->
<!--                  xend = x_to, -->
<!--                  yend = y_to), -->
<!--                colour = "#000000", -->
<!--                alpha = 0.5, -->
<!--                linewidth = 0.5)+ -->
<!--   geom_point(data = umap_two_curvy_scaled, -->
<!--              aes( -->
<!--                x = UMAP1, -->
<!--                y = UMAP2 -->
<!--              ), -->
<!--              color = "#636363", -->
<!--              alpha = 0.5, -->
<!--              size = 0.5 -->
<!--   ) + -->
<!--   coord_equal() + -->
<!--   xlim(c(0, 1)) + -->
<!--   ylim(c(0, 1)) + -->
<!--   interior_annotation("b1", sc_ltr_pos, cex = 2) -->

<!-- ## Benchmark 1 -->

<!-- benchmark_two_curvy <- 2.5 * bin_width -->

<!-- distance_df_small_edges_two_curvy2 <- distance_two_curvy |> -->
<!--   filter(distance < benchmark_two_curvy) ## 0.231 -->

<!-- tr_from_to_df_two_curvy3 <- inner_join( -->
<!--   tr_from_to_df_two_curvy, distance_df_small_edges_two_curvy2, -->
<!--   by = c("from", "to")) -->

<!-- trimesh_two_curvy_removed1_umap <- ggplot() + -->
<!--   geom_segment(data = tr_from_to_df_two_curvy3, -->
<!--                aes( -->
<!--                  x = x_from, -->
<!--                  y = y_from, -->
<!--                  xend = x_to, -->
<!--                  yend = y_to), -->
<!--                colour = "#000000", -->
<!--                alpha = 0.5, -->
<!--                linewidth = 0.5)+ -->
<!--   geom_point(data = umap_two_curvy_scaled, -->
<!--              aes( -->
<!--                x = UMAP1, -->
<!--                y = UMAP2 -->
<!--              ), -->
<!--              color = "#636363", -->
<!--              alpha = 0.5, -->
<!--              size = 0.5 -->
<!--   ) + -->
<!--   coord_equal() + -->
<!--   xlim(c(0, 1)) + -->
<!--   ylim(c(0, 1)) + -->
<!--   interior_annotation("c1", sc_ltr_pos, cex = 2) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- #| label: compare-2d-and-4d-dist -->

<!-- ### Compute highD distance -->
<!-- dist_vec <- proxy::dist(x = df_bin_two_curvy[, -1], method = "Euclidean") |> as.vector() -->

<!-- from_vec <- c() -->
<!-- to_vec <- c() -->
<!-- num_obs <- 1:(NROW(df_bin_two_curvy) - 1) -->

<!-- for (obs in num_obs) { -->

<!--   from_val <- rep(obs, (NROW(df_bin_two_curvy) - obs)) -->
<!--   if ((obs + 1) <= NROW(df_bin_two_curvy)) { -->
<!--     to_val <- (obs + 1):NROW(df_bin_two_curvy) -->
<!--   } -->
<!--   from_vec <- append(from_vec, from_val) -->
<!--   to_vec <- append(to_vec, to_val) -->

<!-- } -->

<!-- dist_highd <- tibble::tibble(from = from_vec, to = to_vec, dist_highd = dist_vec) -->

<!-- ## To plot the distribution of distances -->

<!-- # Define bin width and starting point -->
<!-- start_point <- a1/2 -->

<!-- # Create bins and calculate the mean value for each bin range -->
<!-- bins <- seq(start_point, max(distance_two_curvy$distance) + a1, -->
<!--             by = a1) -->
<!-- bin_labels <- bins[-length(bins)] + a1 / 2  # mean of each bin range -->

<!-- distance_two_curvy_dist <- distance_two_curvy |> -->
<!--   mutate(bin = cut(distance, breaks = bins, include.lowest = TRUE, labels = bin_labels)) -->

<!-- distance_two_curvy_dist <- left_join(distance_two_curvy_dist, dist_highd, by = c("from", "to")) -->

<!-- text_df <- tibble::tibble( -->
<!--   x = c(0.12, 0.43), -->
<!--   y = c(10, 10), -->
<!--   text = c("c", "b") -->
<!-- ) -->

<!-- distance_points_umap <- ggplot( -->
<!--   distance_two_curvy_dist, -->
<!--   aes(x = distance, -->
<!--       y = dist_highd)) + -->
<!--   geom_point(alpha = 0.5) + -->
<!--   geom_text( -->
<!--     data=text_df, -->
<!--     aes(x=x, y=y, -->
<!--         label=text), -->
<!--     colour="#bdbdbd", -->
<!--     size = 10) + -->
<!--   geom_vline(xintercept = 8 * bin_width, -->
<!--              linetype="solid", -->
<!--              color = "#bdbdbd", -->
<!--              linewidth=1) + -->
<!--   geom_vline(xintercept = 2.5 * bin_width, -->
<!--              linetype="solid", -->
<!--              color = "#bdbdbd", -->
<!--              linewidth=1) + -->
<!--   ylab(expression(d^{(4)})) + -->
<!--   xlab(expression(d^{(2)})) + -->
<!--   theme_minimal() + -->
<!--   theme(aspect.ratio = 0.75, -->
<!--         panel.border = element_rect(fill = 'transparent'), -->
<!--         plot.title = element_text(size = 20, hjust = 0.5, vjust = -0.5), -->
<!--         panel.grid.major.x = element_blank(), -->
<!--         axis.ticks.x = element_line(), -->
<!--         axis.ticks.y = element_line(), -->
<!--         axis.title = element_text(size = 15), -->
<!--         axis.text = element_text(size = 12)) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- #| label: proj-tsne-model -->

<!-- df_bin_two_curvy <- df_bin_two_curvy |> -->
<!--   select(-hb_id) |> -->
<!--   mutate(type = "model") -->

<!-- # Apply the scaling -->
<!-- df_model_data_two_curvy <- bind_rows(data_two_curvy, df_bin_two_curvy) -->

<!-- scaled_two_curvy <- scale_data_manual(df_model_data_two_curvy, "type") |> -->
<!--   as_tibble() -->

<!-- scaled_two_curvy_data <- scaled_two_curvy |> -->
<!--   filter(type == "data") |> -->
<!--   select(-type) -->

<!-- scaled_two_curvy_data_model <- scaled_two_curvy |> -->
<!--   filter(type == "model") |> -->
<!--   select(-type) -->

<!-- ## First projection -->
<!-- projection <- cbind( -->
<!--   c(0.37847,-0.06517,0.04231,0.01655), -->
<!--   c(-0.06466,-0.27052,0.22131,-0.15235)) -->

<!-- proj_obj1 <- get_projection(projection = projection,  -->
<!--                             proj_scale = 1,  -->
<!--                             scaled_data = scaled_two_curvy_data,  -->
<!--                             scaled_data_model = scaled_two_curvy_data_model,  -->
<!--                             distance_df_small_edges = distance_df_small_edges_two_curvy2,  -->
<!--                             axis_param = list(limits = 0.6,  -->
<!--                                               axis_scaled = 2, -->
<!--                                               axis_pos_x = -0.5,  -->
<!--                                               axis_pos_y = -0.5,  -->
<!--                                               threshold = 0.04)) -->

<!-- ## Projection with Delauany triangulation -->

<!-- projected_model <- as.matrix(scaled_two_curvy_data_model) %*% projection_scaled -->

<!-- projected_model_df <- projected_model |> -->
<!--   tibble::as_tibble(.name_repair = "unique") |> -->
<!--   dplyr::rename(c("proj1" = "...1", -->
<!--                   "proj2" = "...2")) |> -->
<!--   dplyr::mutate(ID = dplyr::row_number()) -->

<!-- model_df1 <- dplyr::left_join( -->
<!--   distance_two_curvy |> select(-distance), -->
<!--   projected_model_df, -->
<!--   by = c("from" = "ID")) -->

<!-- names(model_df1)[3:NCOL(model_df1)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from") -->

<!-- model_df1 <- dplyr::left_join(model_df1, projected_model_df, by = c("to" = "ID")) -->
<!-- names(model_df1)[(2 + NCOL(projected_model_df)):NCOL(model_df1)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to") -->

<!-- ## With benchmark option 2 -->

<!-- model_df2 <- dplyr::left_join( -->
<!--   distance_df_small_edges_two_curvy3 |> select(-distance), -->
<!--   projected_model_df, -->
<!--   by = c("from" = "ID")) -->

<!-- names(model_df2)[3:NCOL(model_df2)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from") -->

<!-- model_df2 <- dplyr::left_join(model_df2, projected_model_df, by = c("to" = "ID")) -->
<!-- names(model_df2)[(2 + NCOL(projected_model_df)):NCOL(model_df2)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to") -->

<!-- two_curvy_proj_first_model1_umap <- plot_proj( -->
<!--     proj_obj = proj_obj1,  -->
<!--     point_param = c(0.5, 0.2, "#636363"), # size, alpha, color -->
<!--     line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha -->
<!--     plot_limits = c(-0.6, 0.6),  -->
<!--     title = "c2",  -->
<!--     cex = 2,  -->
<!--     position = sc_ltr_pos, -->
<!--     axis_text_size = 4,  -->
<!--     is_color = FALSE) + -->
<!--     theme( -->
<!--         legend.position = "none" -->
<!--     ) -->

<!-- ## Projection with Delauany triangulation -->

<!-- proj_obj1_temp1 <- proj_obj1 -->
<!-- proj_obj1_temp2 <- proj_obj1 -->
<!-- proj_obj1_temp1[["model_df"]] <- model_df1 -->

<!-- two_curvy_proj_model_delaunay_umap <- plot_proj( -->
<!--     proj_obj = proj_obj1_temp1,  -->
<!--     point_param = c(0.5, 0.2, "#636363"), # size, alpha, color -->
<!--     line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha -->
<!--     plot_limits = c(-0.6, 0.6),  -->
<!--     title = "a2",  -->
<!--     cex = 2,  -->
<!--     position = sc_ltr_pos, -->
<!--     axis_text_size = 4,  -->
<!--     is_color = FALSE) + -->
<!--     theme(aspect.ratio = 1,  -->
<!--           legend.position = "none") -->

<!-- proj_obj1_temp2[["model_df"]] <- model_df2 -->
